---
title: "Dataset Creation"
author: "Melvin Coleman"
date: "12/2/2022"
output: html_document
---

```{r setup, include=FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
library(dplyr)
library(readr)
library(rvest)
library(httr)

```

Let's create the dataset we will utilize for the final project. Data was pulled 
from the web via scarping and downloaded from websites. The "data" folder 
contains the csv files of data downloaded and intended for use. 

### World Cup Records & Statistics 
This data was loaded data from Wikipedia and contains record and statistics of
the overall team records of the FIFA World Cup.

```{r, eval=FALSE}
wiki_list =
records_stats_html = read_html("https://en.wikipedia.org/wiki/FIFA_World_Cup_records_and_statistics") %>% 
 html_table(header = TRUE)

wc_stats =
  wiki_list[[3]]
  
wc_stats = 
  wc_stats %>% 
 janitor::clean_names() %>%
   rename(country = team) %>% 
  ## Delete rank(data ranked in order of WC winners) & team
 select(country, everything(), -rank) %>% 
  
  ## Clean dataset, remove weird characters in names 
 mutate(
     country = str_replace(country, "//[c]", "")
  ) %>%
  apply(., 2, function(country) as.character(gsub("\\[|a\\]","",country))) %>%
 apply(., 2, function(country) as.character(gsub("\\[|b\\]","",country))) %>%
  apply(., 2, function(country) as.character(gsub("\\[|c\\]","",country))) %>%
  apply(., 2, function(country) as.character(gsub("\\[|d\\]","",country))) %>%
  apply(., 2, function(country) as.character(gsub("\\[|e\\]","",country))) %>%
  apply(., 2, function(country) as.character(gsub("\\[|f\\]","",country))) %>%
 as_tibble()
  
wc_stats
```


### Fifa Rankings 2022 [Fifa Rankings](https://www.2026worldcupnorthamerica.com/fifa-ranking/)

This dataset was created via web scraping. 
```{r, eval = FALSE}
wc_rank_html =
   read_html("https://www.2026worldcupnorthamerica.com/fifa-ranking/") 

rank_text = 
  wc_rank_html %>% 
  html_elements(".grippy-host , td:nth-child(1), td:nth-child(1)") %>% 
  html_text() 

rank_text

country_text =
  wc_rank_html %>% 
  html_elements("td:nth-child(2)") %>% 
  html_text() 
  
country_text

fifa_rankings = 
tibble(
  rank = rank_text,
  country = country_text
) %>% 
  
  ## Change country names to match other datasets
  mutate(
    country = str_replace(country, "USA","United States"),
    country = str_replace(country, "China PR", "China"),
    country = str_replace(country, "IR Iran", "Iran"),
    country = str_replace(country, "Korea Republic", "South Korea"),
    country = str_replace(country,"Korea DPR","North Korea"),
    country = str_replace(country, "Türkiye", "Turkey"),
    country = str_replace(country,"Czechia", "Czech Republic"),
    country = str_replace(country,"Côte d’Ivoire", "Ivory Coast"),
    country = str_replace(country,  "Congo DR", "DR Congo")
  ) %>% 
  select(country, rank)
  
```

### Confederations Dataset

This dataset was downloaded as a csv file and limited to variables of interest.

```{r, eval=FALSE}
confederations_data = 
  read_csv(file = "data/fifa_countries_audience.csv", col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  ## Select variables of interest 
  select(country, confederation) %>% 
  mutate(
     country = str_replace(country,  "Congo DR", "DR Congo")
  )
  

```

### Top goal scorers per country 

This dataset was pulled from Wikipedia via web scraping. 
```{r, eval=FALSE}
goals_text_html =
   read_html("https://en.wikipedia.org/wiki/List_of_top_international_men%27s_football_goal_scorers_by_country")  %>% 
  html_table(header = TRUE)

goals_country_df = 
  goals_text_html[[1]]


goals_country_df =
  goals_country_df %>% 
  janitor::clean_names() %>% 
  select(country, player, goals)  %>% 
  
## There are countries with more than one goal scorer, let's fix this
  aggregate(player ~ country + goals, FUN = paste, collapse = ' & ') %>% 
  select(country, player, goals) %>% 
  arrange(country) %>% 
  
  ## Change Soviet Union to Russia & update goal scorer info
 mutate(country = str_replace(country,"Soviet Union\\[b]", "Russia"), 
        player = str_replace(player, "Oleg Blokhin", 
        "Alexander Kerzhakov & Artem Dzyuba"))
  
  
  
## Countries to worry about:
## Anguilla, Bulgaria, Curaçao, Denmark,Eritrea, Eswatini, Faroe Islands, France,
## Gibraltar, Iceland, Lebanon, Lesotho, Mongolia, Namibia, Palestine,Romania,
## Scotland, U.S. Virgin Islands, United States

goals_country_df
```

### Population Data 2021 Population 

```{r, eval=FALSE}
pop_df = 
read_csv(file = "data/pop.csv", col_names = TRUE) %>% 
  janitor::clean_names() %>% 
  select(country,land_area_km)
```

Now let's combine all of our datasets to create our final dataset. 

The datasets we have currently are: <br> 
- wc_stats (contains world cup statistics & records)
- fifa_rankings (official Fifa rankings 2022)
- goals_country_df (top goal scorers per country)
- pop_df(land area of countries)
- confederations_data(what confederations each country is in)

We will combine these datasets by the `country` variable, perform some wrangling 
and output final dataset as .csv file. 

```{r, eval=FALSE}
### Put all dataframes into a list & merge by country

merged_df = 
  list(fifa_rankings, goals_country_df, pop_df,confederations_data) 


merged_df2 =
  merged_df %>%
  reduce(full_join, by= "country") %>% 
  arrange(country)
 
final_merge = 
  merge(wc_stats, merged_df2,by= "country")  %>% 
  
  ## Replace missing data with correct info pulled from the web (sources listed below)
  
  mutate(
    ## Add missing confederations for countries 
    confederation = case_when(country %in% c('Wales') ~ 'UEFA', TRUE ~ as.character(confederation)),
    confederation = case_when(country %in% c('Bosnia and Herzegovina') ~ 'UEFA', 
                              TRUE ~ as.character(confederation)),
     confederation = case_when(country %in% c('England') ~ 'UEFA', TRUE ~ as.character(confederation)),
     confederation = case_when(country %in% c('Northern Ireland') ~ 'UEFA',
                              TRUE ~as.character(confederation)),
    confederation = case_when(country %in% c('Republic of Ireland') ~ 'UEFA',
                              TRUE ~as.character(confederation)),
    confederation = case_when(country %in% c('Scotland') ~ 'UEFA',
                              TRUE ~as.character(confederation)),
    confederation = case_when(country %in% c('Trinidad and Tobago') ~ 'CONCACAF',
                              TRUE ~as.character(confederation)),
    confederation = case_when(country %in% c('United Arab Emirates') ~ 'AFC',
                              TRUE ~as.character(confederation)))
```

```{r, eval=FALSE}

## Add missing land area by km squared
final_merge = 
  final_merge %>%
  mutate(
    land_area_km = case_when(country %in% c('Scotland') ~ 77910,
                             TRUE ~ as.numeric(land_area_km)),
    land_area_km = case_when(country %in% c('Republic of Ireland') ~ 70273,
                             TRUE ~ as.numeric(land_area_km)),
    land_area_km = case_when(country %in% c('Northern Ireland') ~ 14130,
                             TRUE ~ as.numeric(land_area_km)),
    land_area_km = case_when(country %in% c('Bosnia and Herzegovina') ~ 51209,
                             TRUE ~ as.numeric(land_area_km)),
    land_area_km = case_when(country %in% c('Trinidad and Tobago') ~ 5128,
                             TRUE ~ as.numeric(land_area_km)),
    land_area_km = case_when(country %in% c('United Arab Emirates') ~ 83600,
                             TRUE ~ as.numeric(land_area_km)), 
    land_area_km = case_when(country %in% c('United Arab Emirates') ~ 83600,
                             TRUE ~ as.numeric(land_area_km)), 
    land_area_km = case_when(country %in% c('England') ~ 130279,
                             TRUE ~ as.numeric(land_area_km)), 
     land_area_km = case_when(country %in% c('Wales') ~ 20780,
                             TRUE ~ as.numeric(land_area_km))
  )
  
```

Output final dataset as `.csv` file for final project use. 

```{r, eval=FALSE}
write.csv(final_merge, "./data/12_4_dataset.csv")
```

```{r, eval=FALSE}
write.csv(final_merge, "./data/world_cup_final.csv")
```

