---
title: "Exploring Predictors of World Cup Wins"
output: 
  html_document:
    toc: TRUE
    toc_float: TRUE
---

<style type="text/css">

h1.title {
  text-align: center;
}

</style>

```{r setup, include=FALSE}
# Load packages
library(tidyverse)
library(viridis)
library(plotly)

# Set ggplot theme
knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

```{r import df, include=FALSE, warning=FALSE, message=FALSE}
# Read in csv datafile
wc_df <- read_csv("data/12_4_dataset.csv") %>%
  janitor::clean_names() %>%
  mutate(
    gd = gsub("\\+","",gd),
    gd = gsub("\\−","-",gd),
    gd = as.numeric(gd)
    ) %>%
  mutate(
    prop_w = w / pld,
    prop_d = d / pld,
    prop_l = l / pld,
    gf_per_game = gf / pld,
    ga_per_game = ga / pld) %>%
  mutate_if(is.numeric, round, digits = 2)
```

<br>

This page contains bivariable explorations between the proportion of games won and various predictors for each country that has participated in the FIFA World Cup. 

<br>

### Codebook

To ensure comparability between countries in this exploratory analysis, the following additional variables were created:

* `prop_w`. Proportion of games a country has won in the World Cup. Calculated as the number of games a country has won in the World Cup divided by the number of games a country has played in the World Cup (`w` / `pld`) <br>
* `gf_per_game`. Average number of goals a country has scored against opponents (i.e. "goals for") per game. Calculated as the total number of goals a country has scored against opponents in the World Cup divided by the number of games a country has played in the World Cup (`gf` / `pld`) <br>
* `ga_per_game`. Average number of goals scored against a country per game. Calculated as the total number of goals scored against a country in the World Cup divided by the number of games a country has played in the World Cup (`ga` / `pld`) <br>

<br>

### 2022 FIFA Ranking

<br>

```{r echo=FALSE, warning=FALSE, message=FALSE}
wc_df %>%
   mutate(text_label = str_c("Country: ", country, "\nConfederation: ", confederation, "\nProportion of Games Won: ", prop_w, "\nFIFA Rank: ", rank)) %>%
  plot_ly(
    x = ~rank, 
    y = ~prop_w,
    type = "scatter",
    mode = "markers",
    color = ~confederation, 
    alpha = 0.7, 
    text = ~text_label, 
    legendgroup = ~confederation,
    colors = viridis_pal(option = "D")(3)) %>% 
  layout(
    title = "2022 FIFA Ranking vs. Proportion of Games Won", 
    xaxis = list(title = "FIFA Ranking (2022)"), 
    yaxis = list(title = "Proportion of Games Won"))
```
There is a negative  association between FIFA Ranking and the proportion of games won in the World Cup. As a team's FIFA Ranking increases (i.e. moves further away from the #1 slot), the proportion of games won in the World Cup decreases. This is expected due to the FIFA ranking scheme that assigns the #1 slot to the top performing team.

<br>

### World Cup Participations

<br>

```{r echo=FALSE, warning=FALSE, message=FALSE}
wc_df %>%
   mutate(text_label = str_c("Country: ", country, "\nConfederation: ", confederation, "\nProportion of Games Won: ", prop_w, "\nWorld Cup Participations: ", part)) %>%
  plot_ly(
    x = ~part, 
    y = ~prop_w,
    type = "scatter",
    mode = "markers",
    color = ~confederation, 
    text = ~text_label, 
    alpha = 0.7, 
    legendgroup = ~confederation,
    colors = viridis_pal(option = "D")(3)) %>% 
  layout(
    title = "Number of World Cup Participations vs. Proportion of Games Won",
    xaxis = list(title = "World Cup Participations"), 
    yaxis = list(title = "Proportion of Games Won"))
```
There is a positive association between number of times a country has participated in the World Cup and the proportion of games won. This is expected, as teams that that win a high proportion of the games are likely some of the best teams and make it to more World Cup tournaments as a result.

<br>

### Games Played

<br>

```{r echo=FALSE, warning=FALSE, message=FALSE}
wc_df %>%
   mutate(text_label = str_c("Country: ", country, "\nConfederation: ", confederation, "\nProportion of Games Won: ", prop_w, "\nNumber of Games Played: ", pld)) %>%
  plot_ly(
    x = ~pld, 
    y = ~prop_w,
    type = "scatter",
    mode = "markers",
    color = ~confederation, 
    text = ~text_label, 
    alpha = 0.7, 
    legendgroup = ~confederation,
    colors = viridis_pal(option = "D")(3)) %>% 
  layout(
    title = "Number of Games Played vs. Proportion of Games Won", 
    xaxis = list(title = "Number of Games Played"), 
    yaxis = list(title = "Proportion of Games Won"))
```
There is a positive association between the number of games played and the proportion of games won in the World Cup. This is expected, as teams that that are doing well and winning a high proportion of games have the opportunity to advance further in the tournament and play a higher number of games.

<br>

### Average Goals For Per Game

<br>

```{r echo=FALSE, warning=FALSE, message=FALSE}
wc_df %>%
   mutate(text_label = str_c("Country: ", country, "\nConfederation: ", confederation, "\nProportion of Games Won: ", prop_w, "\nAvg Goals For Per Game: ", gf_per_game)) %>%
  plot_ly(
    x = ~gf_per_game, 
    y = ~prop_w,
    type = "scatter",
    mode = "markers",
    color = ~confederation, 
    text = ~text_label, 
    alpha = 0.7, 
    legendgroup = ~confederation,
    colors = viridis_pal(option = "D")(3)) %>% 
  layout(
    title = "Average Goals Scored Against Opponent vs. Proportion of Games Won", 
    xaxis = list(title = "Avg Goals For (per game)"), 
    yaxis = list(title = "Proportion of Games Won"))
```
There is a positive association between the average number of goals a country has scored against opponents (i.e. "goals for") per game and the proportion of games a country has won in the World Cup. This is expected, as teams that that are doing well and winning a high proportion of the games are scoring more goals per game than teams that are not doing as well. 

<br>

### Average Goals Against Per Game

<br>

```{r echo=FALSE, warning=FALSE, message=FALSE}
wc_df %>%
   mutate(text_label = str_c("Country: ", country, "\nConfederation: ", confederation, "\nProportion of Games Won: ", prop_w, "\nAvg Goals Against Per Game: ", ga_per_game)) %>%
  plot_ly(
    x = ~ga_per_game, 
    y = ~prop_w,
    type = "scatter",
    mode = "markers",
    color = ~confederation, 
    text = ~text_label, 
    alpha = 0.7, 
    legendgroup = ~confederation,
    colors = viridis_pal(option = "D")(3)) %>% 
  layout(
    title = "Average Goals Scored Against vs. Proportion of Games Won",
    xaxis = list(title = "Avg Goals Against (per game)"), 
    yaxis = list(title = "Proportion of Games Won"))
```
There is a negative association between the average number of goals scored against a country (i.e. "goals against") per game and the proportion of games a country has won in the World Cup. This is expected, as teams that that are doing well and winning a high proportion of the games are receiving less goals scored against them than teams that are not doing as well. 
