---
title: "statistical test"
author: "Jiawen Zhao"
date: "12/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo = FALSE, include=FALSE}
#install.packages('faraway')
#install.packages('Hmisc')
library(Hmisc)
library(faraway)
library(tidyverse)
library(glmnet)
library(MuMIn)
library(corrplot)
library(glmnet)
library(modelr)
library(leaps)
library(caret)
```

Correlation matrix between predictor variables

```{r}
data <- read.csv("./data/12_4_dataset.csv") %>% 
  na.omit()
hist.data.frame(data, nclass = 8)
num = data[3:12] %>% 
  select(-gd)
summary(num)
corrplot(cor(num), type = "upper", diag = FALSE)
```

Model selection: stepwise method


```{r}
model_step_wise = step(glm(gf ~ ., data = data), direction = "both")
model_forward = step(glm(gf ~ 1, data = data), direction = "forward", scope = formula(lm(gf ~ ., data = data)), test = "F")
model_backward = step(glm(gf ~ ., data = data), direction = "backward", test = "F")
model_step_wise
model_forward
model_backward
```

```{r}
predictor_state = num%>%
  select(-gf)

lambda_seq <- 10^seq(-3, 0, by = .1)
set.seed(2020)
cv_object <- cv.glmnet(as.matrix(predictor_state), num$gf,
lambda = lambda_seq,
nfolds = 5)
cv_object
tibble(lambda = cv_object$lambda,
mean_cv_error = cv_object$cvm) %>%
ggplot(aes(x = lambda, y = mean_cv_error)) +
geom_point()
lambda = cv_object$lambda.min

fit_bestcv <- glmnet(as.matrix(predictor_state), num$gf, lambda = cv_object$lambda.min)
coef(fit_bestcv)
lambda

```


Cross-validation of models


```{r}
cv_df = 
  crossv_kfold(data, k = 10) %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  ) %>%
  mutate(
    fit = map(train, ~glm(gf ~ w + ga, data = .x)),
    rmse = map2_dbl(fit, test, ~rmse(model = .x, data = .y))
  )

models <- map(cv_df$train, ~ glm(gf ~ w + ga, data = .))
summary(map2_dbl(models, cv_df$test, modelr::rmse))

cv_df %>%
  ggplot(aes(x = rmse)) +
  geom_density()


cv_df2 = 
  crossv_kfold(data, k = 10) %>%
  mutate(
    train = map(train, as_tibble),
    test = map(test, as_tibble)
  ) %>%
  mutate(
    fit = map(train, ~glm(gf ~ part+l, data = .x)),
    rmse = map2_dbl(fit, test, ~rmse(model = .x, data = .y))
  )

models <- map(cv_df2$train, ~ glm(gf ~ part+l, data = .))
summary(map2_dbl(models, cv_df2$test, modelr::rmse))

cv_df2 %>%
  ggplot(aes(x = rmse)) +
  geom_density()
```


Final regression model output


```{r}
glm(gf ~ w + ga, data = data)
```

Results/output from regression model
With this country's number of goals scored against a country in the World Cup (since 1990) and this country's number of soccer games a country has won in the World Cup (since 1990), we will bw able to predict the number of points to gain given a country in the wourld cup.

This model will be gf = 2.7789gf+0.2573ga.
That means if the number of number of goals scored against a country in the World Cup (since 1990) increases by 1 goal, the would increase by 2.7789.




